libar---
title: "Project Code"
author: "DSW"
date: "2025-03-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r}
library(here)
library(tidyverse)
library(dunn.test)
library(gt)
library(paletteer)
library(moments)
library(dplyr)
library(vcdExtra)
library(ggplot2)
library(readr)
library(dplyr)
library(readxl)
library(leaflet)
library(readr)
library(htmlwidgets)
library(webshot2)
library(ggspatial)
library(sf)
library(maps)
library(viridis)
library(broom)
library(tidyr)
library(e1071)
```

# Wrangling

```{r A horizon}
A_horizon <- read_excel("Data/A_horizon.xls", skip=12)
A_horizon <- A_horizon[-1, ]
View(A_horizon)

# Reference for Units
A_horizon_reference <- read_excel("Data/A_horizon.xls", skip=11)
View(A_horizon_reference)
A_horizon_units <- head(A_horizon_reference, 2)
View(A_horizon_units)



```



# Initial Figures
```{r}
A_states <- A_horizon %>%
  filter(StateID %in% c("NC", "SC", "VA", "GA", "FL"))


# View the filtered data
head(A_states)
```

## Figure 1: Land Cover Frequency by State
```{r Figure 1, fig.cap= "Land status sample type in Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database.}
landcover_counts <- A_states %>%
  group_by(StateID, LandCover1) %>%
  summarize(count = n(), .groups = "drop")

# Create the bar plot
landcover_count <- ggplot(landcover_counts, aes(x = StateID, y = count, fill = LandCover1)) +
  geom_bar(stat = "identity", position = "dodge") +  # 'dodge' makes the bars for each state side by side
  labs(
    title = "Land Cover Frequency by State",
    x = "State",
    y = "Number of Sites",
    fill = "Land Cover Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotates the x-axis labels for better readability
landcover_count
ggsave(landcover_count,
       filename = "landcover_count.jpg",
       height = 12,
       width = 15,
       units = "cm")
```

## Figure 2: (In progress) Total Nutrient Abundance Across States
```{r}
# Convert relevant columns to numeric (direct conversion)
A_states$A_C_Tot <- as.numeric(A_states$A_C_Tot)
A_states$A_Ca <- as.numeric(A_states$A_Ca)
A_states$A_Mg <- as.numeric(A_states$A_Mg)
A_states$A_Tot_K_fs <- as.numeric(A_states$A_Tot_K_fs)  # Use A_Tot_K_fs instead of A_Kaolinit
A_states$A_P <- as.numeric(A_states$A_P)
A_states$A_S <- as.numeric(A_states$A_S)
A_states <- A_states %>%
  mutate(A_P = A_P/10000)

# Now reshape the data into long format for plotting
A_states_long <- A_states %>%
  pivot_longer(cols = c(A_C_Tot, A_Ca, A_Mg, A_Tot_K_fs, A_P, A_S),
               names_to = "Nutrient", values_to = "Abundance")

# Sum the abundances by nutrient
A_states_sum <- A_states_long %>%
  group_by(StateID, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

# Create a bar chart using ggplot
Nut_by_state_bar <-ggplot(A_states_sum, aes(x = StateID, y = Total_Abundance, fill = Nutrient)) +
  geom_bar(stat = "identity", .groups = "drop") +
  labs(title = "Total Macronutrient Abundance Across States",
       x = "State",
       y = "Total Abundance (% weight)",
       fill = "Nutrient") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

Nut_by_state_bar_noP <- ggplot(A_states_sum, aes(x = StateID, y = Total_Abundance, fill = Nutrient)) +
  geom_bar(stat = "identity", .groups = "drop") +
  labs(title = "Total Macronutrient Abundance Across States (exluding P)",
       x = "State",
       y = "Total Abundance (% weight)",
       fill = "Nutrient") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Nut_by_state_bar
Nut_by_state_bar_noP
ggsave(Nut_by_state_bar,
       filename = "nut_by_state_bar.jpg",
       height = 12,
       width = 15,
       units = "cm")
ggsave(Nut_by_state_bar_noP,
       filename = "nut_by_state_bar_noP.jpg",
       height = 12,
       width = 17,
       units = "cm")
```

## Figure 3: Nutrient status by land use type
```{r}
# Examine nutrients and landstatus occurence
 trout_fig3 <- ggplot(A_states, aes(x = A_Ca, fill = LandCover1)) + # base plot
  geom_histogram() + # creates histograms
  scale_fill_manual(values = c("darkorange", "purple", "cyan4","green4","red4", "pink3")) + # customize colors
  labs(x = "Occurence of Ca (% wt)",
       y = "Count") + # edits labels
  facet_grid(.~LandCover1) + # facets by reach to avoid overlap
  theme_bw() + # removes the grey background
  theme(legend.position = "none") # removes the legend

# View figure.
trout_fig3
```

```{r}

#narrowing the dataset to just macro nutrients
A_macros <- A_states %>%
  select(SiteID, StateID, CollDate, LandCover1, LandCover2, A_Depth, A_Tot_K_fs, A_Ca, A_S, A_C_Tot, A_Mg, A_P) %>%
  rename(K = A_Tot_K_fs,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg,
         P = A_P
        )

A_macros_long <- A_macros %>%
  pivot_longer(cols = c(C, Ca, Mg, K, P, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum <- A_macros_long %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

#without P
A_macros_noP <- A_states %>%
  select(SiteID, StateID, CollDate, LandCover1, LandCover2, A_Depth, A_Tot_K_fs, A_Ca, A_S, A_C_Tot, A_Mg) %>%
  rename(K = A_Tot_K_fs,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg)

A_macros_long_noP <- A_macros_noP %>%
  pivot_longer(cols = c(C, Ca, Mg, K, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum_noP <- A_macros_long_noP %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

nutrient_ls_bar <- ggplot(A_macros_sum, aes(x = Nutrient, y = Total_Abundance, fill = Nutrient))+
  geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Total Sum (% Weight)",
       title = "Soil Macronutrient Totals by Land Status Type")+
  scale_fill_viridis(discrete = TRUE)
nutrient_ls_bar
#  caption = "Figure 1: Soil macronutrient sum values based on Land status type only including data from Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database."
nutrient_ls_bar_noP <- ggplot(A_macros_sum_noP, aes(x = Nutrient, y = Total_Abundance, fill = Nutrient))+
  geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Total Sum (% Weight)",
       title = "Soil Macronutrient Totals by Land Status Type (excluding P)")+
  scale_fill_viridis(discrete = TRUE)
nutrient_ls_bar_noP
# caption = "Figure 2: Soil macronutrient sum values, excluding Phosporus to display the lesser occuring macronutrients based on Land status type only including data from Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database."
ggsave(nutrient_ls_bar,
       filename = "nutrient_ls_bar.jpg",
       height = 12,
       width = 15,
       units = "cm")
ggsave(nutrient_ls_bar_noP,
       filename = "nutrient_ls_bar_noP.jpg",
       height = 12,
       width = 15,
       units = "cm")
```


```{r}
trout_fig4 <- ggplot() +
  # add raw data points
  geom_jitter(data = A_states,
              aes(x = LandCover1, y = A_K,
                  color = LandCover1), size = 0.5) +
  # edit colors
  scale_color_manual(values = c("darkorange", "purple", "cyan4", "blue", "lightgreen","red" )) +
  # label axes
  #labs(x = "Reach Section",
       #y = "Snout-to-fork Length (mm)") +
  # remove grey bg
  theme_bw() +
  theme(legend.position = "none")

trout_fig4
```

## Figure 4: Site map
### method (1): leaflet map
```{r}
A_states$Longitude <- as.numeric(A_states$Longitude)
A_states$Latitude <- as.numeric(A_states$Latitude)

pal <- colorFactor(
  palette = c("darkorange", "darkgreen", "cyan4", "blue", "brown","purple"),
  domain = A_states$LandCover1
)

# Create the map
map <- leaflet(A_states) %>%
  # Add base map tiles
  addTiles() %>%
  # Set the initial view to center on your data
  setView(
    lng = mean(A_states$Longitude),
    lat = mean(A_states$Latitude),
    zoom = 5
  ) %>%
  # Add markers for each point
  addCircleMarkers(
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 2,
    #radius = ~value/5,    # Size based on value
    #popup = ~paste("<b>", name, "</b><br>Category: ", category, "<br>Value: ", value),
    #label = ~name,
    color = ~pal(LandCover1),
    #fillOpacity = 0.8,
    stroke = TRUE,
    weight = 1
  ) %>%
  # Add a legend
  addLegend(
    position = "bottomright",
    pal = pal,
    values = ~LandCover1,
    title = "Land Cover Type",
    #opacity = 0.7
  )

# Display the map
map

# First save as HTML
temp_html <- tempfile(fileext = ".html")
saveWidget(map, file = temp_html, selfcontained = TRUE)

# Then convert to PNG
webshot2::webshot(
  url = temp_html,
  file = "site_map.png",
  delay = 5,  # Wait for map to fully load
  zoom = 2    # For higher resolution
)

```

### method (2): ggspatial
```{r}
# Convert to sf object
points_sf <- st_as_sf(A_states, coords = c("Longitude", "Latitude"), crs = 4326)

# Get background map data
usa_map <- map_data("usa")
state_map <- map_data("state")
world_map <- map_data("world")

# Create a static map using ggplot2
map <- ggplot() +
  # Add world base map
  #geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               #fill = "white", color = "gray70", size = 0.2) +
  geom_polygon(data = state_map, aes(x = long, y = lat, group = group),
               fill = NA, color = "gray70", size = 0.3) +
  # Add points with category color and value size
  geom_point(data = A_states, aes(x = Longitude, y = Latitude, color = LandCover1),
             size = 1.2, shape = 16)+
  # Styling
  scale_color_manual(values = c("darkorange", "green", "indianred2", "blue", "pink","purple"),
                     name = "Land Cover Types") +
  #scale_size_continuous(name = "Value", range = c(3, 6)) +
  # Focus map on our data
  coord_sf(xlim = c(min(A_states$Longitude) - 2, max(A_states$Longitude) + 2),
           ylim = c(min(A_states$Latitude) - 1, max(A_states$Latitude) + 1)) +
  # Add map elements
  annotation_north_arrow(location = "tl", style = north_arrow_fancy_orienteering) +
  # Theme and labels
  theme_minimal() +
  labs(title = "Soil Sampling Sites by Land Cover Type",
       caption = "Map showing soil sampling sites (n = 382) in the southeastern U.S. by land cover types",
       x = "Longitude", y = "Latitude") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"), # White background
    plot.background = element_rect(fill = "white", color = "white"), # White plot background
    #panel.grid.major = element_blank(), # Optional: remove gridlines
    #panel.grid.minor = element_blank()  # Optional: remove minor gridlines
  )


print(map)

ggsave("site_map_uniform.png", map, dpi = 300)

```

# Assessing Normality 

## Shapiro Wilk Test

```{r normality shapiro wilk test}
# Shapiro-Wilk test for Carbon (C)
shapiro_C <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "C"])
shapiro_C_tidy <- tidy(shapiro_C)

# Shapiro-Wilk test for Calcium (Ca)
shapiro_Ca <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"])
shapiro_Ca_tidy <- tidy(shapiro_Ca)

# Shapiro-Wilk test for Magnesium (Mg)
shapiro_Mg <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"])
shapiro_Mg_tidy <- tidy(shapiro_Mg)

# Shapiro-Wilk test for Potassium (K)
shapiro_K <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "K"])
shapiro_K_tidy <- tidy(shapiro_K)

# Shapiro-Wilk test for Phosphorus (P)
shapiro_P <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "P"])
shapiro_P_tidy <- tidy(shapiro_P)

# Shapiro-Wilk test for Sulfur (S)
shapiro_S <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "S"])
shapiro_S_tidy <- tidy(shapiro_S)

normality_results <- bind_rows(
  shapiro_C_tidy %>% mutate(Nutrient = "C"),
  shapiro_Ca_tidy %>% mutate(Nutrient = "Ca"),
  shapiro_Mg_tidy %>% mutate(Nutrient = "Mg"),
  shapiro_K_tidy %>% mutate(Nutrient = "K"),
  shapiro_P_tidy %>% mutate(Nutrient = "P"),
  shapiro_S_tidy %>% mutate(Nutrient = "S")
)

print(normality_results)
```
```{r skewness and kurtosis}
abundance_C <- A_macros_long$Abundance[A_macros_long$Nutrient == "C"] %>% na.omit()
if(length(unique(abundance_C)) > 1) {
  skew_C <- skewness(abundance_C)
  kurt_C <- kurtosis(abundance_C)
} else {
  skew_C <- kurt_C <- NA
}

# Skewness and Kurtosis for Calcium (Ca)
abundance_Ca <- A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"] %>% na.omit()
if(length(unique(abundance_Ca)) > 1) {
  skew_Ca <- skewness(abundance_Ca)
  kurt_Ca <- kurtosis(abundance_Ca)
} else {
  skew_Ca <- kurt_Ca <- NA
}

# Skewness and Kurtosis for Magnesium (Mg)
abundance_Mg <- A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"] %>% na.omit()
if(length(unique(abundance_Mg)) > 1) {
  skew_Mg <- skewness(abundance_Mg)
  kurt_Mg <- kurtosis(abundance_Mg)
} else {
  skew_Mg <- kurt_Mg <- NA
}

# Skewness and Kurtosis for Potassium (K)
abundance_K <- A_macros_long$Abundance[A_macros_long$Nutrient == "K"] %>% na.omit()
if(length(unique(abundance_K)) > 1) {
  skew_K <- skewness(abundance_K)
  kurt_K <- kurtosis(abundance_K)
} else {
  skew_K <- kurt_K <- NA
}

# Skewness and Kurtosis for Phosphorus (P)
abundance_P <- A_macros_long$Abundance[A_macros_long$Nutrient == "P"] %>% na.omit()
if(length(unique(abundance_P)) > 1) {
  skew_P <- skewness(abundance_P)
  kurt_P <- kurtosis(abundance_P)
} else {
  skew_P <- kurt_P <- NA
}

# Skewness and Kurtosis for Sulfur (S)
abundance_S <- A_macros_long$Abundance[A_macros_long$Nutrient == "S"] %>% na.omit()
if(length(unique(abundance_S)) > 1) {
  skew_S <- skewness(abundance_S)
  kurt_S <- kurtosis(abundance_S)
} else {
  skew_S <- kurt_S <- NA
}

skew_kurt_results <- data.frame(
  Nutrient = c("C", "Ca", "Mg", "K", "P", "S"),
  Skewness = c(skew_C, skew_Ca, skew_Mg, skew_K, skew_P, skew_S),
  Kurtosis = c(kurt_C, kurt_Ca, kurt_Mg, kurt_K, kurt_P, kurt_S)
)



```

## Normality Histograms

```{r normality plots}

create_histogram <- function(nutrient_name) {
  # Filter data for the selected nutrient
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit()

  # creating the histogram plot
  hist_plot <- ggplot(nutrient_data, aes(x = Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  # Adjust binwidth as needed
    labs(title = paste("Distribution of", nutrient_name, "Abundance by LandCover1"),
         x = paste(nutrient_name, "Abundance"),
         y = "Frequency",
         caption = paste("Figure: Distribution of", nutrient_name, "Abundance by LandCover1")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1)  # Faceting by LandCover1 for each plot

  return(hist_plot)
}


plot_C <- create_histogram("C")
plot_Ca <- create_histogram("Ca")
plot_Mg <- create_histogram("Mg")
plot_K <- create_histogram("K")
plot_P <- create_histogram("P")
plot_S <- create_histogram("S")

# Print each plot
print(plot_C)
print(plot_Ca)
print(plot_Mg)
print(plot_K)
print(plot_P)
print(plot_S)

```

## Log Transformed Normality

```{r Log Transformed Normality}

create_log_transformed_histogram <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1)) 

  hist_plot <- ggplot(nutrient_data, aes(x = log_Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  
    labs(title = paste("Log-Transformed Distribution of", nutrient_name, "Abundance by LandCover1"),
         x = paste("Log-transformed", nutrient_name, "Abundance"),
         y = "Frequency",
         caption = paste("Figure: Log-transformed Distribution of", nutrient_name, "Abundance by LandCover1")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1) 

  return(hist_plot)
}


plot_log_C <- create_log_transformed_histogram("C")
plot_log_Ca <- create_log_transformed_histogram("Ca")
plot_log_Mg <- create_log_transformed_histogram("Mg")
plot_log_K <- create_log_transformed_histogram("K")
plot_log_P <- create_log_transformed_histogram("P")
plot_log_S <- create_log_transformed_histogram("S")

print(plot_log_C)
print(plot_log_Ca)
print(plot_log_Mg)
print(plot_log_K)
print(plot_log_P)
print(plot_log_S)

```

```{r Shapiro Test after log}

shapiro_test_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1))  

  shapiro_result <- shapiro.test(nutrient_data$log_Abundance)

  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_log <- bind_rows(
  shapiro_test_log("C"),
  shapiro_test_log("Ca"),
  shapiro_test_log("Mg"),
  shapiro_test_log("K"),
  shapiro_test_log("P"),
  shapiro_test_log("S")
)

print(shapiro_results_log)

```

```{r shapiro double log}

shapiro_test_double_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1),  # First log transformation
           log_log_Abundance = log(log_Abundance + 1))  # Second log transformation
  
  # Perform the Shapiro-Wilk test on the second log-transformed data
  shapiro_result <- shapiro.test(nutrient_data$log_log_Abundance)

  # Return the result (p-value and W statistic)
  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_double_log <- bind_rows(
  shapiro_test_double_log("C"),
  shapiro_test_double_log("Ca"),
  shapiro_test_double_log("Mg"),
  shapiro_test_double_log("K"),
  shapiro_test_double_log("P"),
  shapiro_test_double_log("S")
)

print(shapiro_results_double_log)

```

















## Linear Regression

## Confidence Interval, T-test
