---
title: "Assessing Influence of Land Cover Type on Soil Nutrient Status in Southeastern U.S."
author: "Daniel Whitehead, Elizabeth Weyl, and P. Louis Lu"
date: "2025-03-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r Packages}
library(here)
library(tidyverse)
library(dunn.test)
library(gt)
library(paletteer)
library(moments)
library(dplyr)
library(vcdExtra)
library(ggplot2)
library(readr)
library(dplyr)
library(readxl)
library(leaflet)
library(readr)
library(htmlwidgets)
library(webshot2)
library(ggspatial)
library(sf)
library(maps)
library(viridis)
library(broom)
library(tidyr)
library(e1071)
library(prism)
library(terra)
library(sp)
library(knitr)
```

# Wrangling

```{r A horizon}
A_horizon <- read_excel("Data/A_horizon.xls", skip=12)
A_horizon <- A_horizon[-1, ]
View(A_horizon)

# Reference for Units
A_horizon_reference <- read_excel("Data/A_horizon.xls", skip=11)
View(A_horizon_reference)
A_horizon_units <- head(A_horizon_reference, 2)
View(A_horizon_units)



```



# Initial Data
```{r Initial Data}
A_states <- A_horizon %>%
  filter(StateID %in% c("NC", "SC", "VA", "GA", "FL"))


# View the filtered data
head(A_states)
```

## Figure 1: Land Cover Frequency by State
```{r Figure 1, fig.cap= "Land status sample type in Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database"}
landcover_counts <- A_states %>%
  group_by(StateID, LandCover1) %>%
  summarize(count = n(), .groups = "drop")

# Create the bar plot
landcover_count <- ggplot(landcover_counts, aes(x = StateID, y = count, fill = LandCover1)) +
  geom_bar(stat = "identity", position = "dodge") +  # 'dodge' makes the bars for each state side by side
  labs(
    title = "Land Cover Frequency by State",
    x = "State",
    y = "Number of Sites",
    fill = "Land Cover Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotates the x-axis labels for better readability
landcover_count
ggsave(landcover_count,
       filename = "landcover_count.jpg",
       height = 12,
       width = 15,
       units = "cm")
```

## Figure 2: Total Nutrient Abundance Across States
```{r Total Nutrient Abundance Across States}
# Convert relevant columns to numeric (direct conversion)
A_states$A_C_Tot <- as.numeric(A_states$A_C_Tot)
A_states$A_Ca <- as.numeric(A_states$A_Ca)
A_states$A_Mg <- as.numeric(A_states$A_Mg)
#A_states$A_Tot_K_fs <- as.numeric(A_states$A_Tot_K_fs) # Use A_Tot_K_fs instead of A_Kaolinit
A_states$A_K <- as.numeric(A_states$A_K) # No, we should use A_K instead of A_Tot_K_fs
A_states$A_P <- as.numeric(A_states$A_P)
A_states$A_S <- as.numeric(A_states$A_S)
A_states <- A_states %>%
  mutate(A_P = A_P/10000)

# Now reshape the data into long format for plotting
A_states_long <- A_states %>%
  pivot_longer(cols = c(A_C_Tot, A_Ca, A_Mg, A_K, A_P, A_S),
               names_to = "Nutrient", values_to = "Abundance")

# Sum the abundances by nutrient
A_states_sum <- A_states_long %>%
  group_by(StateID, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))
```

# Original bar plots 
```{r Original Bar Plots}

#narrowing the dataset to just macro nutrients
A_macros <- A_states %>%
  select(SiteID, StateID, CollDate, Longitude, Latitude, LandCover1, LandCover2, A_Depth, A_K, A_Ca, A_S, A_C_Tot, A_Mg, A_P, A_Tot_Clay) %>%
  rename(K = A_K,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg,
         P = A_P
        )

A_macros_long <- A_macros %>%
  pivot_longer(cols = c(C, Ca, Mg, K, P, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum <- A_macros_long %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

#without P
A_macros_noP <- A_states %>%
  select(SiteID, StateID, CollDate, LandCover1, LandCover2, A_Depth, A_K, A_Ca, A_S, A_C_Tot, A_Mg) %>%
  rename(K = A_K,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg)

A_macros_long_noP <- A_macros_noP %>%
  pivot_longer(cols = c(C, Ca, Mg, K, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum_noP <- A_macros_long_noP %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))


```

# Macronutrients boxplots by land cover type
```{r Boxplots}
A_macros_long <- A_macros_long %>%
  filter(LandCover1 != "Non-Natural Woody" & 
         LandCover1 != "Shrubland")

macronutrient_box <- ggplot(A_macros_long, aes(x = Nutrient, y = Abundance, fill = Nutrient))+
  geom_boxplot() +
  #geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Abundance (% Weight)",
       title = "Distribution of Soil Macronutrients by Primary Land Cover") +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_log10()
macronutrient_box

annotations <- data.frame(
  Nutrient = rep(c("C", "Ca", "K", "Mg", "P", "S"), each = 4),  # 3 annotations per facet
  x = c(1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4,
        1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4),  # x positions for each annotation in the respective facet
  y = c(80, 80, 80, 80, 80, 80, 80, 80,
        8, 8, 8, 8, 5, 5, 5, 5,
        0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5),  # y positions for each annotation
  label = c("a", "b", "a", "a", 
            "a", "b", "b", "c",
            "a", "b", "a", "a",
            "a", "b", "b", "a",
            "a", "b", "b", "b",
            "a", "b", "b", "c"))  # Text for each annotation
  
# Boxplots by Nutrients
macronutrient_box1 <- ggplot(A_macros_long, aes(x = LandCover1, y = Abundance, fill = Nutrient))+
  geom_boxplot() +
  #geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(Nutrient), nrow = 2, scales = "free_y")+
  theme_bw()+ 
  labs(x = "Land Cover Type",
       y = "Abundance (% Weight)",
       title = "Distribution of Soil Macronutrients by Primary Land Cover") +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_log10() +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Developed" = "Developed", 
                              "Forested Upland" = "Forested",
                              "Herbaceous Upland" = "Herbaceous",
                              "Planted/Cultivated" = "Cultivated")) +
  geom_text(data = annotations, aes(x = x, y = y, label = label), inherit.aes = FALSE, size = 4, color = "red")
macronutrient_box1

ggsave("macronutrient_boxplot_annotated.png", plot = macronutrient_box1, width = 10, height = 6, units = "in")

ggsave("macronutrient_boxplot.png", plot = macronutrient_box, width = 10, height = 6, units = "in")
```


## Figure 4: Site map

```{r Site Map}
# Make long & lat numeric
A_states$Longitude <- as.numeric(A_states$Longitude)
A_states$Latitude <- as.numeric(A_states$Latitude)
# Convert to sf object
points_sf <- st_as_sf(A_states, coords = c("Longitude", "Latitude"), crs = 4326)

# Get background map data
usa_map <- map_data("usa")
state_map <- map_data("state")
world_map <- map_data("world")

# Create a static map using ggplot2
map <- ggplot() +
  # Add world base map
  #geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               #fill = "white", color = "gray70", size = 0.2) +
  geom_polygon(data = state_map, aes(x = long, y = lat, group = group),
               fill = NA, color = "gray70", size = 0.3) +
  # Add points with category color and value size
  geom_point(data = A_states, aes(x = Longitude, y = Latitude, color = LandCover1),
             size = 1.2, shape = 16)+
  # Styling
  scale_color_manual(values = c("darkorange", "green", "indianred2", "blue", "pink","purple"),
                     name = "Land Cover Types") +
  #scale_size_continuous(name = "Value", range = c(3, 6)) +
  # Focus map on our data
  coord_sf(xlim = c(min(A_states$Longitude) - 2, max(A_states$Longitude) + 2),
           ylim = c(min(A_states$Latitude) - 1, max(A_states$Latitude) + 1)) +
  # Add map elements
  annotation_north_arrow(location = "tl", style = north_arrow_fancy_orienteering) +
  # Theme and labels
  theme_minimal() +
  labs(title = "Soil Sampling Sites by Land Cover Type",
       caption = "Map showing soil sampling sites (n = 382) in the southeastern U.S. by land cover types",
       x = "Longitude", y = "Latitude") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"), # White background
    plot.background = element_rect(fill = "white", color = "white"), # White plot background
    #panel.grid.major = element_blank(), # Optional: remove gridlines
    #panel.grid.minor = element_blank()  # Optional: remove minor gridlines
  )


print(map)

ggsave("site_map_uniform.png", map, dpi = 300)

```

# Confidence Interval

```{r Confidence Interval Prep}
# filter the dataset by four land cover types
forest_macro <- A_macros %>%
  filter(LandCover1 == "Forested Upland")

developed_macro <- A_macros %>%
  filter(LandCover1 == "Developed")

herbaceous_macro <- A_macros %>%
  filter(LandCover1 == "Herbaceous Upland")

cultivated_macro <- A_macros %>%
  filter(LandCover1 == "Planted/Cultivated")

A_macros_4 <- A_macros %>%
  filter(LandCover1 != "Non-Natural Woody" & 
         LandCover1 != "Shrubland")
```

### Calculate 95% CI for each macronutrient by land cover
```{r 95% CI}
# Carbon
forest_C_test <- t.test(forest_macro$C, conf.level = 0.95)

forest_C_test

developed_C_test <- t.test(developed_macro$C, conf.level = 0.95)

developed_C_test

herbaceous_C_test <- t.test(herbaceous_macro$C, conf.level = 0.95)

herbaceous_C_test

cultivated_C_test <- t.test(cultivated_macro$C, conf.level = 0.95)

cultivated_C_test

C_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_C = mean(C, na.rm = TRUE),
                       sd = sd(C, na.rm = TRUE),
                       CI95_lower = t.test(C, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(C, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
C_summary
```

```{r Calcium ttest}
# Ca
forest_Ca_test <- t.test(forest_macro$Ca, conf.level = 0.95)

forest_Ca_test

developed_Ca_test <- t.test(developed_macro$Ca, conf.level = 0.95)

developed_Ca_test

herbaceous_Ca_test <- t.test(herbaceous_macro$Ca, conf.level = 0.95)

herbaceous_Ca_test

cultivated_Ca_test <- t.test(cultivated_macro$Ca, conf.level = 0.95)

cultivated_Ca_test

Ca_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_S = mean(Ca, na.rm = TRUE),
                       sd = sd(Ca, na.rm = TRUE),
                       CI95_lower = t.test(Ca, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(Ca, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
Ca_summary
```

```{r K ttest}
# K
forest_K_test <- t.test(forest_macro$K, conf.level = 0.95)

forest_K_test

developed_K_test <- t.test(developed_macro$K, conf.level = 0.95)

developed_K_test

herbaceous_K_test <- t.test(herbaceous_macro$K, conf.level = 0.95)

herbaceous_K_test

cultivated_K_test <- t.test(cultivated_macro$K, conf.level = 0.95)

cultivated_K_test

K_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_K = mean(K, na.rm = TRUE),
                       sd = sd(K, na.rm = TRUE),
                       CI95_lower = t.test(K, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(K, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
K_summary
```

```{r Mg Ttest}
# Mg
forest_Mg_test <- t.test(forest_macro$Mg, conf.level = 0.95)

forest_Mg_test

developed_Mg_test <- t.test(developed_macro$Mg, conf.level = 0.95)

developed_Mg_test

herbaceous_Mg_test <- t.test(herbaceous_macro$Mg, conf.level = 0.95)

herbaceous_Mg_test

cultivated_Mg_test <- t.test(cultivated_macro$Mg, conf.level = 0.95)

cultivated_Mg_test

Mg_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_Mg = mean(Mg, na.rm = TRUE),
                       sd = sd(Mg, na.rm = TRUE),
                       CI95_lower = t.test(Mg, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(Mg, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
Mg_summary
```

```{r P ttest}
# P
forest_P_test <- t.test(forest_macro$P, conf.level = 0.95)

forest_P_test

developed_P_test <- t.test(developed_macro$P, conf.level = 0.95)

developed_P_test

herbaceous_P_test <- t.test(herbaceous_macro$P, conf.level = 0.95)

herbaceous_P_test

cultivated_P_test <- t.test(cultivated_macro$P, conf.level = 0.95)

cultivated_P_test

P_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_P = mean(P, na.rm = TRUE),
                       sd = sd(P, na.rm = TRUE),
                       CI95_lower = t.test(P, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(P, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
P_summary
```

```{r S ttest}
# S
forest_S_test <- t.test(forest_macro$S, conf.level = 0.95)

forest_S_test

developed_S_test <- t.test(developed_macro$S, conf.level = 0.95)

developed_S_test

herbaceous_S_test <- t.test(herbaceous_macro$S, conf.level = 0.95)

herbaceous_S_test

cultivated_S_test <- t.test(cultivated_macro$S, conf.level = 0.95)

cultivated_S_test

S_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_S = mean(S, na.rm = TRUE),
                       sd = sd(S, na.rm = TRUE),
                       CI95_lower = t.test(S, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(S, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
S_summary
```



# Assessing Normality 

## Shapiro Wilk Test

```{r normality shapiro wilk test}
# Shapiro-Wilk test for Carbon (C)
shapiro_C <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "C"])
shapiro_C_tidy <- tidy(shapiro_C)

# Shapiro-Wilk test for Calcium (Ca)
shapiro_Ca <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"])
shapiro_Ca_tidy <- tidy(shapiro_Ca)

# Shapiro-Wilk test for Magnesium (Mg)
shapiro_Mg <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"])
shapiro_Mg_tidy <- tidy(shapiro_Mg)

# Shapiro-Wilk test for Potassium (K)
shapiro_K <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "K"])
shapiro_K_tidy <- tidy(shapiro_K)

# Shapiro-Wilk test for Phosphorus (P)
shapiro_P <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "P"])
shapiro_P_tidy <- tidy(shapiro_P)

# Shapiro-Wilk test for Sulfur (S)
shapiro_S <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "S"])
shapiro_S_tidy <- tidy(shapiro_S)

normality_results <- bind_rows(
  shapiro_C_tidy %>% mutate(Nutrient = "C"),
  shapiro_Ca_tidy %>% mutate(Nutrient = "Ca"),
  shapiro_Mg_tidy %>% mutate(Nutrient = "Mg"),
  shapiro_K_tidy %>% mutate(Nutrient = "K"),
  shapiro_P_tidy %>% mutate(Nutrient = "P"),
  shapiro_S_tidy %>% mutate(Nutrient = "S")
)

print(normality_results)
```
## Skewness and kurtosis

```{r skewness and kurtosis}
abundance_C <- A_macros_long$Abundance[A_macros_long$Nutrient == "C"] %>% na.omit()
if(length(unique(abundance_C)) > 1) {
  skew_C <- skewness(abundance_C)
  kurt_C <- kurtosis(abundance_C)
} else {
  skew_C <- kurt_C <- NA
}

# Skewness and Kurtosis for Calcium (Ca)
abundance_Ca <- A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"] %>% na.omit()
if(length(unique(abundance_Ca)) > 1) {
  skew_Ca <- skewness(abundance_Ca)
  kurt_Ca <- kurtosis(abundance_Ca)
} else {
  skew_Ca <- kurt_Ca <- NA
}

# Skewness and Kurtosis for Magnesium (Mg)
abundance_Mg <- A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"] %>% na.omit()
if(length(unique(abundance_Mg)) > 1) {
  skew_Mg <- skewness(abundance_Mg)
  kurt_Mg <- kurtosis(abundance_Mg)
} else {
  skew_Mg <- kurt_Mg <- NA
}

# Skewness and Kurtosis for Potassium (K)
abundance_K <- A_macros_long$Abundance[A_macros_long$Nutrient == "K"] %>% na.omit()
if(length(unique(abundance_K)) > 1) {
  skew_K <- skewness(abundance_K)
  kurt_K <- kurtosis(abundance_K)
} else {
  skew_K <- kurt_K <- NA
}

# Skewness and Kurtosis for Phosphorus (P)
abundance_P <- A_macros_long$Abundance[A_macros_long$Nutrient == "P"] %>% na.omit()
if(length(unique(abundance_P)) > 1) {
  skew_P <- skewness(abundance_P)
  kurt_P <- kurtosis(abundance_P)
} else {
  skew_P <- kurt_P <- NA
}

# Skewness and Kurtosis for Sulfur (S)
abundance_S <- A_macros_long$Abundance[A_macros_long$Nutrient == "S"] %>% na.omit()
if(length(unique(abundance_S)) > 1) {
  skew_S <- skewness(abundance_S)
  kurt_S <- kurtosis(abundance_S)
} else {
  skew_S <- kurt_S <- NA
}

skew_kurt_results <- data.frame(
  Nutrient = c("C", "Ca", "Mg", "K", "P", "S"),
  Skewness = c(skew_C, skew_Ca, skew_Mg, skew_K, skew_P, skew_S),
  Kurtosis = c(kurt_C, kurt_Ca, kurt_Mg, kurt_K, kurt_P, kurt_S)
)



```

## Normality Histograms

```{r normality plots}

create_histogram <- function(nutrient_name) {
  # Filter data for the selected nutrient
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit()

  # creating the histogram plot
  hist_plot <- ggplot(nutrient_data, aes(x = Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  # Adjust binwidth as needed
    labs(title = paste("Distribution of", nutrient_name, "Abundance by Primary Land Cover"),
         x = paste(nutrient_name, "Abundance by % weight"),
         y = "Frequency",
         caption = paste("Figure: Distribution of", nutrient_name, "Abundance by Primary Land Cover")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1)  # Faceting by LandCover1 for each plot

  return(hist_plot)
}


plot_C <- create_histogram("C")
plot_Ca <- create_histogram("Ca")
plot_Mg <- create_histogram("Mg")
plot_K <- create_histogram("K")
plot_P <- create_histogram("P")
plot_S <- create_histogram("S")

# Print each plot
print(plot_C)
print(plot_Ca)
print(plot_Mg)
print(plot_K)
print(plot_P)
print(plot_S)

```

## Log Transformed Normality

```{r Log Transformed Normality}

create_log_transformed_histogram <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1)) 

  hist_plot <- ggplot(nutrient_data, aes(x = log_Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  
    labs(title = paste("Log-Transformed Distribution of", nutrient_name, "Abundance by LandCover1"),
         x = paste("Log-transformed", nutrient_name, "Abundance by % weight"),
         y = "Frequency",
         caption = paste("Figure: Log-transformed Distribution of", nutrient_name, "Abundance by LandCover1")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1) 

  return(hist_plot)
}


plot_log_C <- create_log_transformed_histogram("C")
plot_log_Ca <- create_log_transformed_histogram("Ca")
plot_log_Mg <- create_log_transformed_histogram("Mg")
plot_log_K <- create_log_transformed_histogram("K")
plot_log_P <- create_log_transformed_histogram("P")
plot_log_S <- create_log_transformed_histogram("S")

print(plot_log_C)
print(plot_log_Ca)
print(plot_log_Mg)
print(plot_log_K)
print(plot_log_P)
print(plot_log_S)

```

```{r Shapiro Test after log}

shapiro_test_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1))  

  shapiro_result <- shapiro.test(nutrient_data$log_Abundance)

  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_log <- bind_rows(
  shapiro_test_log("C"),
  shapiro_test_log("Ca"),
  shapiro_test_log("Mg"),
  shapiro_test_log("K"),
  shapiro_test_log("P"),
  shapiro_test_log("S")
)

print(shapiro_results_log)
shapiro_results_log1 <- shapiro_results_log %>%
  select(Nutrient, p_value) # Modify this depending on your actual column names


kable(shapiro_results_log1, col.names = c("Nutrient", "P-Value"), caption = "Shapiro-Wilk Test Results for Nutrients")
```

```{r shapiro double log}

shapiro_test_double_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1),  # First log transformation
           log_log_Abundance = log(log_Abundance + 1))  # Second log transformation
  
  # Perform the Shapiro-Wilk test on the second log-transformed data
  shapiro_result <- shapiro.test(nutrient_data$log_log_Abundance)

  # Return the result (p-value and W statistic)
  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_double_log <- bind_rows(
  shapiro_test_double_log("C"),
  shapiro_test_double_log("Ca"),
  shapiro_test_double_log("Mg"),
  shapiro_test_double_log("K"),
  shapiro_test_double_log("P"),
  shapiro_test_double_log("S")
)

print(shapiro_results_double_log)

```

## Kruskal Wallis and Dunn Tests

#### Kruskal-Wallis and Dunn's Test for Nutrient C

```{r Kruskal-Wallis and Dunn Test for Nutrient C}
nutrient_C <- A_macros_long %>%
  filter(Nutrient == "C") %>%
  na.omit()

# Run Kruskal-Wallis test for C
kruskal_C <- kruskal.test(Abundance ~ LandCover1, data = nutrient_C)
print("Kruskal-Wallis Test Result for C:")
print(kruskal_C)

dunn_C <- dunn.test(nutrient_C$Abundance, nutrient_C$LandCover1, kw = TRUE, label = TRUE)
print(dunn_C)

```

### Kruskal-Wallis and Dunn's Test for Nutrient Ca

```{r Kruskal-Wallis and Dunn Test for Nutrient Ca}
nutrient_Ca <- A_macros_long %>%
  filter(Nutrient == "Ca") %>%
  na.omit()

# Run Kruskal-Wallis test for Ca
kruskal_Ca <- kruskal.test(Abundance ~ LandCover1, data = nutrient_Ca)
print("Kruskal-Wallis Test Result for Ca:")
print(kruskal_Ca)

dunn_Ca <- dunn.test(nutrient_Ca$Abundance, nutrient_Ca$LandCover1, kw = TRUE, label = TRUE)
print(dunn_Ca)

```

### Kruskal-Wallis and Dunn's Test for Nutrient Mg

```{r Kruskal-Wallis and Dunn Test for Nutrient Mg}
nutrient_Mg <- A_macros_long %>%
  filter(Nutrient == "Mg") %>%
  na.omit()

# Run Kruskal-Wallis test for Mg
kruskal_Mg <- kruskal.test(Abundance ~ LandCover1, data = nutrient_Mg)
print("Kruskal-Wallis Test Result for Mg:")
print(kruskal_Mg)


dunn_Mg <- dunn.test(nutrient_Mg$Abundance, nutrient_Mg$LandCover1, kw = TRUE, label = TRUE)
print(dunn_Mg)

```

### Kruskal-Wallis and Dunn's Test for Nutrient K

```{r Kruskal-Wallis and Dunn Test for Nutrient K}
nutrient_K <- A_macros_long %>%
  filter(Nutrient == "K") %>%
  na.omit()

# Run Kruskal-Wallis test for K
kruskal_K <- kruskal.test(Abundance ~ LandCover1, data = nutrient_K)
print("Kruskal-Wallis Test Result for K:")
print(kruskal_K)

dunn_K <- dunn.test(nutrient_K$Abundance, nutrient_K$LandCover1, kw = TRUE, label = TRUE)
print(dunn_K)

```

### Kruskal-Wallis and Dunn's Test for Nutrient P

```{r Kruskal-Wallis and Dunn Test for Nutrient P}
nutrient_P <- A_macros_long %>%
  filter(Nutrient == "P") %>%
  na.omit()

# Run Kruskal-Wallis test for P
kruskal_P <- kruskal.test(Abundance ~ LandCover1, data = nutrient_P)
print("Kruskal-Wallis Test Result for P:")
print(kruskal_P)

dunn_P <- dunn.test(nutrient_P$Abundance, nutrient_P$LandCover1, kw = TRUE, label = TRUE)
print(dunn_P)

```

### Kruskal-Wallis and Dunn's Test for Nutrient S

```{r Kruskal-Wallis and Dunn Test for Nutrient S}
nutrient_S <- A_macros_long %>%
  filter(Nutrient == "S") %>%
  na.omit()

# Run Kruskal-Wallis test for S
kruskal_S <- kruskal.test(Abundance ~ LandCover1, data = nutrient_S)
print("Kruskal-Wallis Test Result for S:")
print(kruskal_S)

dunn_S <- dunn.test(nutrient_S$Abundance, nutrient_S$LandCover1, kw = TRUE, label = TRUE)
print(dunn_S)


```






## Linear Regression

## Mean temperature and Precipitation as potential independent variables
```{r Temp and Precipitation Data}
# Set download directory
prism_set_dl_dir('./Data')
# Download temperature data for a specific region
get_prism_normals(type = "tmean", 
                resolution = "4km",  # 4km spatial resolution (choose "4km" or "800m")
                #mon = 1:12,
                annual = TRUE,
                keepZip = FALSE)

get_prism_normals(type = "ppt", 
                resolution = "4km",  # 4km spatial resolution (choose "4km" or "800m")
                #mon = 1:12,
                annual = TRUE,
                keepZip = FALSE)
```

```{r Rasterize Data}
prism_t <- rast("./Data/PRISM_tmean_30yr_normal_4kmM5_annual_bil/PRISM_tmean_30yr_normal_4kmM5_annual_bil.bil")
prism_p <- rast("./Data/PRISM_ppt_30yr_normal_4kmM4_annual_bil/PRISM_ppt_30yr_normal_4kmM4_annual_bil.bil")
```

```{r Plot temp and precipiation}
temp_plot<- plot(prism_t,
     main = "Average Annual Temperature °C",
     xlab = "Longitude",
     ylab = "Latitude")
precipitation_plot <- plot(prism_p,
     main = "Annual Precipitation (mm)",
     xlab = "Longitude",
     ylab = "Latitude")

```

```{r Extracting positions}
A_macros$Longitude <- as.numeric(A_macros$Longitude)
A_macros$Latitude <- as.numeric(A_macros$Latitude)

# Define the coordinates
coords <- A_macros %>%
  select(Longitude, Latitude)
 # rename(x = Longitude, y = Latitude)

coords <- data.frame(coords)

points <- vect(coords, geom=c("Longitude", "Latitude"), crs = "EPSG:4326")
print(points)

points_nad83 <- project(points, crs(prism_t))

print(points_nad83)

extracted_t <- extract(prism_t, points_nad83)
extracted_p <- extract(prism_p, points_nad83)

A_macros_tp <- A_macros %>%
  mutate(temperature = extracted_t,
         precipitation = extracted_p) %>%
  rename(clay = A_Tot_Clay,
         depth = A_Depth)
```

## Multiple Linear Regression
## Carbon

```{r tidying exogenous variables}
# convert Depth to average depth
A_macros_tp <- A_macros_tp %>%
  mutate(avg_depth = sapply(strsplit(as.character(depth), "-"), 
                            function(x) mean(as.numeric(x)))) %>%
# tidy clay percentage data
  mutate(clay = na_if(clay, "N.D."),
         clay = as.numeric(clay)) %>%
  filter(LandCover1 != "Non-Natural Woody" & 
         LandCover1 != "Shrubland")

A_macros_tp_C <- A_macros_tp %>%
  select(C, 
         LandCover1,
         clay, 
         avg_depth)

A_macros_tp_C$temperature <- A_macros_tp$temperature$PRISM_tmean_30yr_normal_4kmM5_annual_bil
A_macros_tp_C$precipitation <- A_macros_tp$precipitation$PRISM_ppt_30yr_normal_4kmM4_annual_bil
```

### View correlation
```{r Viewing Correlation}
# Pivot all independent variables but not the dependent variable.
A_macros_tp_C_long <- A_macros_tp_C %>%
  select(C, 
         clay, 
         avg_depth,
         temperature,
         precipitation) %>%
pivot_longer(names_to = "variable", cols = -C, values_to = "value")

# Create and examine an initial scatterplot of data.
ggplot(A_macros_tp_C_long, aes(x = value, y = C, color = variable)) +
geom_point() +
facet_wrap(~variable, scales = "free_x") +
theme_minimal() +
theme(legend.position = "none") +
labs(title = "Scatterplots of Each Independent Var v.s. Dependent Var",
x = "Independent Variable (Name in Subtitles)",
y = "Total Recreation Acreage")

# Correlation matrix
A_macros_tp_C1 <- A_macros_tp_C %>%
  select(C, clay, avg_depth, temperature, precipitation) %>%
  na.omit()
cor_matrix_C <- cor(A_macros_tp_C1)

View(cor_matrix_C)

```

#### Model Building
```{r Model Building}
# Set Landcover column to be a factor with a reference level of Forested.
A_macros_tp_C$LandCover1 <- factor(A_macros_tp_C$LandCover1,
                             levels = c("Forested Upland", "Developed", 
                                        "Planted/Cultivated", "Herbaceous Upland"))
```


```{r Carbon MLM1}
C.mlm1 <- lm(C ~ clay + avg_depth +
               temperature + precipitation +
               LandCover1,
             data = A_macros_tp_C)

summary(C.mlm1)

plot(C.mlm1)
```

#### log transformed data

```{r Log Transformed Temp and Precip Data}
A_macros_tp_C_log <- A_macros_tp_C %>%
  mutate(C_log = log(C),
         clay_log = log(clay),
         temp_log = log(temperature),
         precip_log = log(precipitation),
         depth_log = log(avg_depth))
```

#### Refit with log-transformed data

```{r Refitting with log data}
C.mlm2 <- lm(C_log ~ clay_log + depth_log +
               temp_log + precip_log +
               LandCover1,
             data = A_macros_tp_C_log)

summary(C.mlm2)

plot(C.mlm2)
```

## Linear Model visualization

```{r LM Visualization}
# Simulate temp and precip data
temp_vector <- rep(seq(from = 10, to = 30), 4)
precip_vector <-rep(seq(from = 800, to = 2000, length.out = 21), 4)
clay_vector <-rep(seq(from = 0.1, to = 40, length.out = 21), 4)
depth_vector <-rep(seq(from = 5, to = 35, length.out = 21), 4)

temp_vector_log = log(temp_vector)
precip_vector_log = log(precip_vector)
clay_vector_log = log(clay_vector)
depth_vector_log = log(depth_vector)

# Simulate landcover data.
land_vector <- c(rep("Forested Upland", 21),
                 rep("Developed", 21),
                 rep("Planted/Cultivated", 21),
                 rep("Herbaceous Upland", 21))

# Join with continuous data.
data_pred <- data.frame(temp_vector_log, precip_vector_log, clay_vector_log, depth_vector_log, land_vector)
colnames(data_pred) <- c("temp_log", "precip_log", "clay_log", "depth_log", "LandCover1")

# Use the original model to predict outcomes.
prediction <- predict(C.mlm2,
newdata = data_pred,
type = "response",
se.fit = TRUE)

# Pull out predictions.
data_fig <- data.frame(data_pred,
prediction$fit,
prediction$se.fit)

# Rename columns.
colnames(data_fig) <- c("temp_log", "precip_log", "clay_log", "depth_log", "LandCover1", "probability", "se")

# Graph the probabilities of getting a whole tail.
model_C_fig <- ggplot(data_fig, aes(x = depth_log,
                                    y = probability)) +
               geom_line(aes(color = LandCover1)) +
               geom_ribbon(aes(ymin = probability - se,
                               ymax = probability + se,
                               fill = LandCover1), alpha = 0.3) +
               labs(title = "Effects of Land Cover Types on Soil Carbon Content",
                    x = "Log Average Depth", y = "Predicted Log C Content",
                    color = "LandCover1", fill = "LandCover1") +
               theme_bw()

ggsave("Carbon_model.png", plot = model_C_fig, width = 8, height = 6)
```



