libar---
title: "Project Code"
author: "DSW"
date: "2025-03-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r}
library(here)
library(tidyverse)
library(dunn.test)
library(gt)
library(paletteer)
library(moments)
library(dplyr)
library(vcdExtra)
library(ggplot2)
library(readr)
library(dplyr)
library(readxl)
library(leaflet)
library(readr)
library(htmlwidgets)
library(webshot2)
library(ggspatial)
library(sf)
library(maps)
library(viridis)
library(broom)
library(tidyr)
library(e1071)
```

# Wrangling

```{r A horizon}
A_horizon <- read_excel("Data/A_horizon.xls", skip=12)
A_horizon <- A_horizon[-1, ]
View(A_horizon)

# Reference for Units
A_horizon_reference <- read_excel("Data/A_horizon.xls", skip=11)
View(A_horizon_reference)
A_horizon_units <- head(A_horizon_reference, 2)
View(A_horizon_units)



```



# Initial Figures
```{r}
A_states <- A_horizon %>%
  filter(StateID %in% c("NC", "SC", "VA", "GA", "FL"))


# View the filtered data
head(A_states)
```

## Figure 1: Land Cover Frequency by State
```{r Figure 1, fig.cap= "Land status sample type in Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database.}
landcover_counts <- A_states %>%
  group_by(StateID, LandCover1) %>%
  summarize(count = n(), .groups = "drop")

# Create the bar plot
landcover_count <- ggplot(landcover_counts, aes(x = StateID, y = count, fill = LandCover1)) +
  geom_bar(stat = "identity", position = "dodge") +  # 'dodge' makes the bars for each state side by side
  labs(
    title = "Land Cover Frequency by State",
    x = "State",
    y = "Number of Sites",
    fill = "Land Cover Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotates the x-axis labels for better readability
landcover_count
ggsave(landcover_count,
       filename = "landcover_count.jpg",
       height = 12,
       width = 15,
       units = "cm")
```

## Figure 2: (In progress) Total Nutrient Abundance Across States
```{r}
# Convert relevant columns to numeric (direct conversion)
A_states$A_C_Tot <- as.numeric(A_states$A_C_Tot)
A_states$A_Ca <- as.numeric(A_states$A_Ca)
A_states$A_Mg <- as.numeric(A_states$A_Mg)
#A_states$A_Tot_K_fs <- as.numeric(A_states$A_Tot_K_fs) # Use A_Tot_K_fs instead of A_Kaolinit
A_states$A_K <- as.numeric(A_states$A_K) # No, we should use A_K instead of A_Tot_K_fs
A_states$A_P <- as.numeric(A_states$A_P)
A_states$A_S <- as.numeric(A_states$A_S)
A_states <- A_states %>%
  mutate(A_P = A_P/10000)

# Now reshape the data into long format for plotting
A_states_long <- A_states %>%
  pivot_longer(cols = c(A_C_Tot, A_Ca, A_Mg, A_K, A_P, A_S),
               names_to = "Nutrient", values_to = "Abundance")

# Sum the abundances by nutrient
A_states_sum <- A_states_long %>%
  group_by(StateID, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

# Create a bar chart using ggplot
Nut_by_state_bar <-ggplot(A_states_sum, aes(x = StateID, y = Total_Abundance, fill = Nutrient)) +
  geom_bar(stat = "identity", .groups = "drop") +
  labs(title = "Total Macronutrient Abundance Across States",
       x = "State",
       y = "Total Abundance (% weight)",
       fill = "Nutrient") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

Nut_by_state_bar_noP <- ggplot(A_states_sum, aes(x = StateID, y = Total_Abundance, fill = Nutrient)) +
  geom_bar(stat = "identity", .groups = "drop") +
  labs(title = "Total Macronutrient Abundance Across States (exluding P)",
       x = "State",
       y = "Total Abundance (% weight)",
       fill = "Nutrient") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Nut_by_state_bar
Nut_by_state_bar_noP
ggsave(Nut_by_state_bar,
       filename = "nut_by_state_bar.jpg",
       height = 12,
       width = 15,
       units = "cm")
ggsave(Nut_by_state_bar_noP,
       filename = "nut_by_state_bar_noP.jpg",
       height = 12,
       width = 17,
       units = "cm")
```

# Original bar plots (Now I think the boxplots underneath look better)
```{r}

#narrowing the dataset to just macro nutrients
A_macros <- A_states %>%
  select(SiteID, StateID, CollDate, LandCover1, LandCover2, A_Depth, A_K, A_Ca, A_S, A_C_Tot, A_Mg, A_P) %>%
  rename(K = A_K,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg,
         P = A_P
        )

A_macros_long <- A_macros %>%
  pivot_longer(cols = c(C, Ca, Mg, K, P, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum <- A_macros_long %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

#without P
A_macros_noP <- A_states %>%
  select(SiteID, StateID, CollDate, LandCover1, LandCover2, A_Depth, A_K, A_Ca, A_S, A_C_Tot, A_Mg) %>%
  rename(K = A_K,
         Ca = A_Ca,
         S = A_S,
         C = A_C_Tot,
         Mg = A_Mg)

A_macros_long_noP <- A_macros_noP %>%
  pivot_longer(cols = c(C, Ca, Mg, K, S),
               names_to = "Nutrient", values_to = "Abundance")

A_macros_sum_noP <- A_macros_long_noP %>%
  group_by(StateID, LandCover1, Nutrient) %>%
  summarize(Total_Abundance = sum(Abundance, na.rm = TRUE))

nutrient_ls_bar <- ggplot(A_macros_sum, aes(x = Nutrient, y = Total_Abundance, fill = Nutrient))+
  geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Total Sum (% Weight)",
       title = "Soil Macronutrient Totals by Land Status Type")+
  scale_fill_viridis(discrete = TRUE)
nutrient_ls_bar
#  caption = "Figure 1: Soil macronutrient sum values based on Land status type only including data from Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database."
nutrient_ls_bar_noP <- ggplot(A_macros_sum_noP, aes(x = Nutrient, y = Total_Abundance, fill = Nutrient))+
  geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Total Sum (% Weight)",
       title = "Soil Macronutrient Totals by Land Status Type (excluding P)")+
  scale_fill_viridis(discrete = TRUE)
nutrient_ls_bar_noP
# caption = "Figure 2: Soil macronutrient sum values, excluding Phosporus to display the lesser occuring macronutrients based on Land status type only including data from Florida, Georgia, North Carolina, South Carolina, and Virginia. This data is from the USGS Geochemical and mineralogical data for soils of the conterminous United States database."
ggsave(nutrient_ls_bar,
       filename = "nutrient_ls_bar.jpg",
       height = 12,
       width = 15,
       units = "cm")
ggsave(nutrient_ls_bar_noP,
       filename = "nutrient_ls_bar_noP.jpg",
       height = 12,
       width = 15,
       units = "cm")
```

# Macronutrients boxplots by land cover type
```{r}
A_macros_long <- A_macros_long %>%
  filter(LandCover1 != "Non-Natural Woody" & 
         LandCover1 != "Shrubland")

macronutrient_box <- ggplot(A_macros_long, aes(x = Nutrient, y = Abundance, fill = Nutrient))+
  geom_boxplot() +
  #geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(LandCover1), nrow = 2)+
  theme_bw()+ 
  labs(x = "Nutrient",
       y = "Abundance (% Weight)",
       title = "Distribution of Soil Macronutrients by Land Cover Type") +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_log10()
macronutrient_box

# Boxplots by Nutrients
macronutrient_box1 <- ggplot(A_macros_long, aes(x = LandCover1, y = Abundance, fill = Nutrient))+
  geom_boxplot() +
  #geom_bar(stat = "identity", .groups = "drop")+
  facet_wrap(vars(Nutrient), nrow = 2, scales = "free_y")+
  theme_bw()+ 
  labs(x = "Land Cover Type",
       y = "Abundance (% Weight)",
       title = "Distribution of Soil Macronutrients by Land Cover Type") +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_log10() +
  theme(legend.position = "none")
macronutrient_box1
```

## Figure 4: Site map

```{r}
# Make long & lat numeric
A_states$Longitude <- as.numeric(A_states$Longitude)
A_states$Latitude <- as.numeric(A_states$Latitude)
# Convert to sf object
points_sf <- st_as_sf(A_states, coords = c("Longitude", "Latitude"), crs = 4326)

# Get background map data
usa_map <- map_data("usa")
state_map <- map_data("state")
world_map <- map_data("world")

# Create a static map using ggplot2
map <- ggplot() +
  # Add world base map
  #geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               #fill = "white", color = "gray70", size = 0.2) +
  geom_polygon(data = state_map, aes(x = long, y = lat, group = group),
               fill = NA, color = "gray70", size = 0.3) +
  # Add points with category color and value size
  geom_point(data = A_states, aes(x = Longitude, y = Latitude, color = LandCover1),
             size = 1.2, shape = 16)+
  # Styling
  scale_color_manual(values = c("darkorange", "green", "indianred2", "blue", "pink","purple"),
                     name = "Land Cover Types") +
  #scale_size_continuous(name = "Value", range = c(3, 6)) +
  # Focus map on our data
  coord_sf(xlim = c(min(A_states$Longitude) - 2, max(A_states$Longitude) + 2),
           ylim = c(min(A_states$Latitude) - 1, max(A_states$Latitude) + 1)) +
  # Add map elements
  annotation_north_arrow(location = "tl", style = north_arrow_fancy_orienteering) +
  # Theme and labels
  theme_minimal() +
  labs(title = "Soil Sampling Sites by Land Cover Type",
       caption = "Map showing soil sampling sites (n = 382) in the southeastern U.S. by land cover types",
       x = "Longitude", y = "Latitude") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"), # White background
    plot.background = element_rect(fill = "white", color = "white"), # White plot background
    #panel.grid.major = element_blank(), # Optional: remove gridlines
    #panel.grid.minor = element_blank()  # Optional: remove minor gridlines
  )


print(map)

ggsave("site_map_uniform.png", map, dpi = 300)

```

# Confidence Interval

```{r}
# filter the dataset by four land cover types
forest_macro <- A_macros %>%
  filter(LandCover1 == "Forested Upland")

developed_macro <- A_macros %>%
  filter(LandCover1 == "Developed")

herbaceous_macro <- A_macros %>%
  filter(LandCover1 == "Herbaceous Upland")

cultivated_macro <- A_macros %>%
  filter(LandCover1 == "Planted/Cultivated")

A_macros_4 <- A_macros %>%
  filter(LandCover1 != "Non-Natural Woody" & 
         LandCover1 != "Shrubland")
```

### Calculate 95% CI for each macronutrient by land cover
```{r}
# Carbon
forest_C_test <- t.test(forest_macro$C, conf.level = 0.95)

forest_C_test

developed_C_test <- t.test(developed_macro$C, conf.level = 0.95)

developed_C_test

herbaceous_C_test <- t.test(herbaceous_macro$C, conf.level = 0.95)

herbaceous_C_test

cultivated_C_test <- t.test(cultivated_macro$C, conf.level = 0.95)

cultivated_C_test

C_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_C = mean(C, na.rm = TRUE),
                       sd = sd(C, na.rm = TRUE),
                       CI95_lower = t.test(C, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(C, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
C_summary
```

```{r}
# Ca
forest_Ca_test <- t.test(forest_macro$Ca, conf.level = 0.95)

forest_Ca_test

developed_Ca_test <- t.test(developed_macro$Ca, conf.level = 0.95)

developed_Ca_test

herbaceous_Ca_test <- t.test(herbaceous_macro$Ca, conf.level = 0.95)

herbaceous_Ca_test

cultivated_Ca_test <- t.test(cultivated_macro$Ca, conf.level = 0.95)

cultivated_Ca_test

Ca_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_S = mean(Ca, na.rm = TRUE),
                       sd = sd(Ca, na.rm = TRUE),
                       CI95_lower = t.test(Ca, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(Ca, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
Ca_summary
```

```{r}
# K
forest_K_test <- t.test(forest_macro$K, conf.level = 0.95)

forest_K_test

developed_K_test <- t.test(developed_macro$K, conf.level = 0.95)

developed_K_test

herbaceous_K_test <- t.test(herbaceous_macro$K, conf.level = 0.95)

herbaceous_K_test

cultivated_K_test <- t.test(cultivated_macro$K, conf.level = 0.95)

cultivated_K_test

K_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_K = mean(K, na.rm = TRUE),
                       sd = sd(K, na.rm = TRUE),
                       CI95_lower = t.test(K, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(K, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
K_summary
```

```{r}
# Mg
forest_Mg_test <- t.test(forest_macro$Mg, conf.level = 0.95)

forest_Mg_test

developed_Mg_test <- t.test(developed_macro$Mg, conf.level = 0.95)

developed_Mg_test

herbaceous_Mg_test <- t.test(herbaceous_macro$Mg, conf.level = 0.95)

herbaceous_Mg_test

cultivated_Mg_test <- t.test(cultivated_macro$Mg, conf.level = 0.95)

cultivated_Mg_test

Mg_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_Mg = mean(Mg, na.rm = TRUE),
                       sd = sd(Mg, na.rm = TRUE),
                       CI95_lower = t.test(Mg, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(Mg, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
Mg_summary
```

```{r}
# P
forest_P_test <- t.test(forest_macro$P, conf.level = 0.95)

forest_P_test

developed_P_test <- t.test(developed_macro$P, conf.level = 0.95)

developed_P_test

herbaceous_P_test <- t.test(herbaceous_macro$P, conf.level = 0.95)

herbaceous_P_test

cultivated_P_test <- t.test(cultivated_macro$P, conf.level = 0.95)

cultivated_P_test

P_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_P = mean(P, na.rm = TRUE),
                       sd = sd(P, na.rm = TRUE),
                       CI95_lower = t.test(P, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(P, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
P_summary
```

```{r}
# S
forest_S_test <- t.test(forest_macro$S, conf.level = 0.95)

forest_S_test

developed_S_test <- t.test(developed_macro$S, conf.level = 0.95)

developed_S_test

herbaceous_S_test <- t.test(herbaceous_macro$S, conf.level = 0.95)

herbaceous_S_test

cultivated_S_test <- t.test(cultivated_macro$S, conf.level = 0.95)

cultivated_S_test

S_summary <- A_macros_4 %>%
             group_by(LandCover1) %>%
             summarize(mean_S = mean(S, na.rm = TRUE),
                       sd = sd(S, na.rm = TRUE),
                       CI95_lower = t.test(S, conf.level = 0.95)$conf.int[1],
                       CI95_upper = t.test(S, conf.level = 0.95)$conf.int[2]) %>%
             ungroup()

# View
S_summary
```



# Assessing Normality 

## Shapiro Wilk Test

```{r normality shapiro wilk test}
# Shapiro-Wilk test for Carbon (C)
shapiro_C <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "C"])
shapiro_C_tidy <- tidy(shapiro_C)

# Shapiro-Wilk test for Calcium (Ca)
shapiro_Ca <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"])
shapiro_Ca_tidy <- tidy(shapiro_Ca)

# Shapiro-Wilk test for Magnesium (Mg)
shapiro_Mg <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"])
shapiro_Mg_tidy <- tidy(shapiro_Mg)

# Shapiro-Wilk test for Potassium (K)
shapiro_K <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "K"])
shapiro_K_tidy <- tidy(shapiro_K)

# Shapiro-Wilk test for Phosphorus (P)
shapiro_P <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "P"])
shapiro_P_tidy <- tidy(shapiro_P)

# Shapiro-Wilk test for Sulfur (S)
shapiro_S <- shapiro.test(A_macros_long$Abundance[A_macros_long$Nutrient == "S"])
shapiro_S_tidy <- tidy(shapiro_S)

normality_results <- bind_rows(
  shapiro_C_tidy %>% mutate(Nutrient = "C"),
  shapiro_Ca_tidy %>% mutate(Nutrient = "Ca"),
  shapiro_Mg_tidy %>% mutate(Nutrient = "Mg"),
  shapiro_K_tidy %>% mutate(Nutrient = "K"),
  shapiro_P_tidy %>% mutate(Nutrient = "P"),
  shapiro_S_tidy %>% mutate(Nutrient = "S")
)

print(normality_results)
```
```{r skewness and kurtosis}
abundance_C <- A_macros_long$Abundance[A_macros_long$Nutrient == "C"] %>% na.omit()
if(length(unique(abundance_C)) > 1) {
  skew_C <- skewness(abundance_C)
  kurt_C <- kurtosis(abundance_C)
} else {
  skew_C <- kurt_C <- NA
}

# Skewness and Kurtosis for Calcium (Ca)
abundance_Ca <- A_macros_long$Abundance[A_macros_long$Nutrient == "Ca"] %>% na.omit()
if(length(unique(abundance_Ca)) > 1) {
  skew_Ca <- skewness(abundance_Ca)
  kurt_Ca <- kurtosis(abundance_Ca)
} else {
  skew_Ca <- kurt_Ca <- NA
}

# Skewness and Kurtosis for Magnesium (Mg)
abundance_Mg <- A_macros_long$Abundance[A_macros_long$Nutrient == "Mg"] %>% na.omit()
if(length(unique(abundance_Mg)) > 1) {
  skew_Mg <- skewness(abundance_Mg)
  kurt_Mg <- kurtosis(abundance_Mg)
} else {
  skew_Mg <- kurt_Mg <- NA
}

# Skewness and Kurtosis for Potassium (K)
abundance_K <- A_macros_long$Abundance[A_macros_long$Nutrient == "K"] %>% na.omit()
if(length(unique(abundance_K)) > 1) {
  skew_K <- skewness(abundance_K)
  kurt_K <- kurtosis(abundance_K)
} else {
  skew_K <- kurt_K <- NA
}

# Skewness and Kurtosis for Phosphorus (P)
abundance_P <- A_macros_long$Abundance[A_macros_long$Nutrient == "P"] %>% na.omit()
if(length(unique(abundance_P)) > 1) {
  skew_P <- skewness(abundance_P)
  kurt_P <- kurtosis(abundance_P)
} else {
  skew_P <- kurt_P <- NA
}

# Skewness and Kurtosis for Sulfur (S)
abundance_S <- A_macros_long$Abundance[A_macros_long$Nutrient == "S"] %>% na.omit()
if(length(unique(abundance_S)) > 1) {
  skew_S <- skewness(abundance_S)
  kurt_S <- kurtosis(abundance_S)
} else {
  skew_S <- kurt_S <- NA
}

skew_kurt_results <- data.frame(
  Nutrient = c("C", "Ca", "Mg", "K", "P", "S"),
  Skewness = c(skew_C, skew_Ca, skew_Mg, skew_K, skew_P, skew_S),
  Kurtosis = c(kurt_C, kurt_Ca, kurt_Mg, kurt_K, kurt_P, kurt_S)
)



```

## Normality Histograms

```{r normality plots}

create_histogram <- function(nutrient_name) {
  # Filter data for the selected nutrient
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit()

  # creating the histogram plot
  hist_plot <- ggplot(nutrient_data, aes(x = Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  # Adjust binwidth as needed
    labs(title = paste("Distribution of", nutrient_name, "Abundance by LandCover1"),
         x = paste(nutrient_name, "Abundance"),
         y = "Frequency",
         caption = paste("Figure: Distribution of", nutrient_name, "Abundance by LandCover1")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1)  # Faceting by LandCover1 for each plot

  return(hist_plot)
}


plot_C <- create_histogram("C")
plot_Ca <- create_histogram("Ca")
plot_Mg <- create_histogram("Mg")
plot_K <- create_histogram("K")
plot_P <- create_histogram("P")
plot_S <- create_histogram("S")

# Print each plot
print(plot_C)
print(plot_Ca)
print(plot_Mg)
print(plot_K)
print(plot_P)
print(plot_S)

```

## Log Transformed Normality

```{r Log Transformed Normality}

create_log_transformed_histogram <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1)) 

  hist_plot <- ggplot(nutrient_data, aes(x = log_Abundance, fill = LandCover1)) +
    geom_histogram(binwidth = 0.1, color = "black", position = "dodge") +  
    labs(title = paste("Log-Transformed Distribution of", nutrient_name, "Abundance by LandCover1"),
         x = paste("Log-transformed", nutrient_name, "Abundance"),
         y = "Frequency",
         caption = paste("Figure: Log-transformed Distribution of", nutrient_name, "Abundance by LandCover1")) +
    scale_fill_manual(values = c("LandCover1_Value1" = "skyblue", "LandCover1_Value2" = "salmon")) +  # Replace with actual values for LandCover1
    theme_minimal() +
    theme(legend.position = "right") + 
    facet_grid(. ~ LandCover1) 

  return(hist_plot)
}


plot_log_C <- create_log_transformed_histogram("C")
plot_log_Ca <- create_log_transformed_histogram("Ca")
plot_log_Mg <- create_log_transformed_histogram("Mg")
plot_log_K <- create_log_transformed_histogram("K")
plot_log_P <- create_log_transformed_histogram("P")
plot_log_S <- create_log_transformed_histogram("S")

print(plot_log_C)
print(plot_log_Ca)
print(plot_log_Mg)
print(plot_log_K)
print(plot_log_P)
print(plot_log_S)

```

```{r Shapiro Test after log}

shapiro_test_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1))  

  shapiro_result <- shapiro.test(nutrient_data$log_Abundance)

  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_log <- bind_rows(
  shapiro_test_log("C"),
  shapiro_test_log("Ca"),
  shapiro_test_log("Mg"),
  shapiro_test_log("K"),
  shapiro_test_log("P"),
  shapiro_test_log("S")
)

print(shapiro_results_log)

```

```{r shapiro double log}

shapiro_test_double_log <- function(nutrient_name) {
  nutrient_data <- A_macros_long %>%
    filter(Nutrient == nutrient_name) %>%
    na.omit() %>%
    mutate(log_Abundance = log(Abundance + 1),  # First log transformation
           log_log_Abundance = log(log_Abundance + 1))  # Second log transformation
  
  # Perform the Shapiro-Wilk test on the second log-transformed data
  shapiro_result <- shapiro.test(nutrient_data$log_log_Abundance)

  # Return the result (p-value and W statistic)
  return(data.frame(Nutrient = nutrient_name, 
                    W_statistic = shapiro_result$statistic, 
                    p_value = shapiro_result$p.value))
}

shapiro_results_double_log <- bind_rows(
  shapiro_test_double_log("C"),
  shapiro_test_double_log("Ca"),
  shapiro_test_double_log("Mg"),
  shapiro_test_double_log("K"),
  shapiro_test_double_log("P"),
  shapiro_test_double_log("S")
)

print(shapiro_results_double_log)

```

## Anova and Tukey

```{r}


# Run ANOVA for "C"
nutrient_C <- A_macros_long %>%
  filter(Nutrient == "C") %>%
  na.omit()

anova_C <- aov(Abundance ~ LandCover1, data = nutrient_C)
summary_C <- summary(anova_C)
print("ANOVA Result for C:")
print(summary_C)

# Perform Tukey HSD for "C"
tukey_C <- TukeyHSD(anova_C)
print("Tukey HSD for C:")
print(tukey_C)

# Run ANOVA for "Ca"
nutrient_Ca <- A_macros_long %>%
  filter(Nutrient == "Ca") %>%
  na.omit()

anova_Ca <- aov(Abundance ~ LandCover1, data = nutrient_Ca)
summary_Ca <- summary(anova_Ca)
print("ANOVA Result for Ca:")
print(summary_Ca)

# Perform Tukey HSD for "Ca"
tukey_Ca <- TukeyHSD(anova_Ca)
print("Tukey HSD for Ca:")
print(tukey_Ca)

# Run ANOVA for "Mg"
nutrient_Mg <- A_macros_long %>%
  filter(Nutrient == "Mg") %>%
  na.omit()

anova_Mg <- aov(Abundance ~ LandCover1, data = nutrient_Mg)
summary_Mg <- summary(anova_Mg)
print("ANOVA Result for Mg:")
print(summary_Mg)

# Perform Tukey HSD for "Mg"
tukey_Mg <- TukeyHSD(anova_Mg)
print("Tukey HSD for Mg:")
print(tukey_Mg)

# Run ANOVA for "K"
nutrient_K <- A_macros_long %>%
  filter(Nutrient == "K") %>%
  na.omit()

anova_K <- aov(Abundance ~ LandCover1, data = nutrient_K)
summary_K <- summary(anova_K)
print("ANOVA Result for K:")
print(summary_K)

# Perform Tukey HSD for "K"
tukey_K <- TukeyHSD(anova_K)
print("Tukey HSD for K:")
print(tukey_K)

# Run ANOVA for "P"
nutrient_P <- A_macros_long %>%
  filter(Nutrient == "P") %>%
  na.omit()

anova_P <- aov(Abundance ~ LandCover1, data = nutrient_P)
summary_P <- summary(anova_P)
print("ANOVA Result for P:")
print(summary_P)

# Perform Tukey HSD for "P"
tukey_P <- TukeyHSD(anova_P)
print("Tukey HSD for P:")
print(tukey_P)

# Run ANOVA for "S"
nutrient_S <- A_macros_long %>%
  filter(Nutrient == "S") %>%
  na.omit()

anova_S <- aov(Abundance ~ LandCover1, data = nutrient_S)
summary_S <- summary(anova_S)
print("ANOVA Result for S:")
print(summary_S)

# Perform Tukey HSD for "S"
tukey_S <- TukeyHSD(anova_S)
print("Tukey HSD for S:")
print(tukey_S)


```

## Kruskal Wallis and Dunn Tests

#### Kruskal-Wallis and Dunn's Test for Nutrient C

```{r}
nutrient_C <- A_macros_long %>%
  filter(Nutrient == "C") %>%
  na.omit()

# Run Kruskal-Wallis test for C
kruskal_C <- kruskal.test(Abundance ~ LandCover1, data = nutrient_C)
print("Kruskal-Wallis Test Result for C:")
print(kruskal_C)

dunn_C <- dunn.test(nutrient_C$Abundance, nutrient_C$LandCover1, kw = TRUE, label = TRUE)
print(dunn_C)

```

### Kruskal-Wallis and Dunn's Test for Nutrient Ca

```{r}
nutrient_Ca <- A_macros_long %>%
  filter(Nutrient == "Ca") %>%
  na.omit()

# Run Kruskal-Wallis test for Ca
kruskal_Ca <- kruskal.test(Abundance ~ LandCover1, data = nutrient_Ca)
print("Kruskal-Wallis Test Result for Ca:")
print(kruskal_Ca)

dunn_Ca <- dunn.test(nutrient_Ca$Abundance, nutrient_Ca$LandCover1, kw = TRUE, label = TRUE)
print(dunn_Ca)

```

### Kruskal-Wallis and Dunn's Test for Nutrient Mg

```{r}
nutrient_Mg <- A_macros_long %>%
  filter(Nutrient == "Mg") %>%
  na.omit()

# Run Kruskal-Wallis test for Mg
kruskal_Mg <- kruskal.test(Abundance ~ LandCover1, data = nutrient_Mg)
print("Kruskal-Wallis Test Result for Mg:")
print(kruskal_Mg)


dunn_Mg <- dunn.test(nutrient_Mg$Abundance, nutrient_Mg$LandCover1, kw = TRUE, label = TRUE)
print(dunn_Mg)

```

### Kruskal-Wallis and Dunn's Test for Nutrient K

```{r}
nutrient_K <- A_macros_long %>%
  filter(Nutrient == "K") %>%
  na.omit()

# Run Kruskal-Wallis test for K
kruskal_K <- kruskal.test(Abundance ~ LandCover1, data = nutrient_K)
print("Kruskal-Wallis Test Result for K:")
print(kruskal_K)

dunn_K <- dunn.test(nutrient_K$Abundance, nutrient_K$LandCover1, kw = TRUE, label = TRUE)
print(dunn_K)

```

### Kruskal-Wallis and Dunn's Test for Nutrient P

```{r}
nutrient_P <- A_macros_long %>%
  filter(Nutrient == "P") %>%
  na.omit()

# Run Kruskal-Wallis test for P
kruskal_P <- kruskal.test(Abundance ~ LandCover1, data = nutrient_P)
print("Kruskal-Wallis Test Result for P:")
print(kruskal_P)

dunn_P <- dunn.test(nutrient_P$Abundance, nutrient_P$LandCover1, kw = TRUE, label = TRUE)
print(dunn_P)

```

### Kruskal-Wallis and Dunn's Test for Nutrient S

```{r}
nutrient_S <- A_macros_long %>%
  filter(Nutrient == "S") %>%
  na.omit()

# Run Kruskal-Wallis test for S
kruskal_S <- kruskal.test(Abundance ~ LandCover1, data = nutrient_S)
print("Kruskal-Wallis Test Result for S:")
print(kruskal_S)

dunn_S <- dunn.test(nutrient_S$Abundance, nutrient_S$LandCover1, kw = TRUE, label = TRUE)
print(dunn_S)


```






## Linear Regression

## Confidence Interval


